General constraints on magic distillation protocols from statistical mechanics

### 2021-05-26 magic_paper.tex
small changes

If we have a magic state $\rho$ that has negativity in its Wigner representation, then, as discussed, its Lorenz curve over-shoots $1$ and reaches a non-trivial maximum $L_\star$ that depends on the particular state. There is an exact relation between $L_\star$ and sum-negativity or mana, which is provided by the following result. 
\begin{theorem}\label{lem:lcmax}
	Given a magic state $\rho$, the maximum $L_\star$ of its Lorenz curve $\lc{\rho}{\sigma}(x)$ is independent of the $\sigma$--fragment and equal to $1+\sn{\rho}$. Moreover, the majorization constraint is stronger than mana in every fragment.
\end{theorem}
The proof of this is given in~\cref{app:major}. Using the Lorenz curve perspective, it is simple to construct other monotones. For example, within a given fragment, the area above $L=1$ is another magic monotone.
\begin{lemma}
Given a magic state $\rho$ and a free state $\sigma$, let $\A_\sigma(\rho)$ be the area of the region $\{(x, y): 1 \leq y \leq L_{\rho | \sigma}(x)\}$. Then $\A_\sigma(\rho)$ is a magic monotone for $\R_\sigma$.
\end{lemma}
\begin{proof}
Consider the transformation $\rho_1 \rightarrow \rho_2$ within $\R_\sigma$. We have that $L_{\rho_1|\sigma}(x)$ is never below the curve $L_{\rho_2|\sigma}(x)$, and therefore the region above $L=1$ for $\rho_2$ is a subset of the corresponding region for $\rho_1$. Thus, $\A_\sigma(\rho_1) \geq \A_\sigma(\rho_2)$, so $\A_\sigma$ is a magic monotone in $\R_\sigma$.
\end{proof}
In contrast to mana, the area monotone is \emph{specific to the fragment}, and its value will vary as we change $\sigma$. Therefore its monotonicity is also including the physics of the fragment and provides a means to analyse magic distillation only under free operations that leave $\sigma$ invariant.

@@

We now show that the Lorenz curve is concave in the context of quasi-distributions.
\begin{lemma}\label{L-concave} Let $\p$ be a quasi-probability distribution and let $\r$ be a probability distribution with strictly non-zero components. Then $L_{\p|\r}(x)$ is a concave function on $[0,1]$.
\nick{
\begin{proof}
	The Lorenz curve of $\p$ relative to $\r$ passes through $(0,0)$ and the points $(\sum_{i=1}^k{r_{\pi(i)}}, \sum_{i=1}^k p_{\pi(i)})$ where $\pi$ is the permutation that puts $p_i/r_i$ in non-increasing order.
	By construction, the slope of the curve attains the values $p_{\pi(i)}/r_{\pi(i)}$ throughout the interval $[0,1]$, which is a non-increasing sequence due to the definition of $\pi$.
	Therefore, the Lorenz curve is concave.
\end{proof}
}
\begin{proof}Let $L_{\p|\r}(x)$ attain its maximum at $x=x_\star$. Then from the definition, the function is continuous, monotone increasing on $[0,x_\star]$, and monotone decreasing on $[x_\star, 1]$. Consider $a,b \in [0,1]$ with $a <b$ and some probability $q \in [0,1]$. If $a$ and $b$ both lie in either the interval $[0,x_\star]$ or $[x_\star, 1]$ then concavity follows immediately. Otherwise, $a$ is in the first interval and $b$ in the second interval. If $w=qa + (1-q)b$ lies in the first interval then
\begin{align}
qL_{\p|\r}(a) + (1-q)L_{\p|\r}(b) &\le qL_{\p|\r}(a) + (1-q)L_{\p|\r}(x_\star)\nonumber\\
& \le L_{\p|\r}(w)
\end{align}
\nick{second inequality isn't clear to me}
from monotonicity on the first interval. Otherwise, $w$ lies in the second interval and
\begin{align}
qL_{\p|\r}(a) + (1-q)L_{\p|\r}(b) &\le qL_{\p|\r}(x_\star) + (1-q)L_{\p|\r}(b)\nonumber\\
& \le L_{\p|\r}(w)
\end{align}
\nick{again second inequality isn't clear to me}
which follows from monotonicity on the second interval. Therefore the Lorenz curve is concave on $[0,1]$.
\end{proof}
\end{lemma}

@@

In the unital fragment, the rescaled distribution is simply proportional to the Strange state distribution,
\begin{equation}
	W_{\rho_{\rm{S}}(\epsilon)^{\otimes n}|\left(\id/3\right)^{\otimes n}} = 9^n W_{\rho_{\rm{S}}(\epsilon)^{\otimes n}},
\end{equation}
therefore expressions for the Lorenz curve coordinates follow from summing up the Wigner components in decreasing order.
\nick{As a result, every Strange state Lorenz curve in the unital fragment generally contains $n$ elbows. In the case of accidental degeneracies among Wigner components, some elbows my be straightened out.}
Including the boundary points $(x_{-1}, L_{-1}) := (0,0)$ and $(x_{n}, L_{n}) := (1,1)$, we label these elbows as 
\begin{equation*}
\{(x_{i}, L_{i})\}_{i=-1,0,\dots,n},
\end{equation*}
where the coordinates are given by
\begin{equation}
	(x_{i}, L_{i}) = \left( \frac{1}{9^n}\sum_{\ell=0}^i m_{\ell}, \sum_{\ell=0}^i m_{\ell} w_{\ell} \right).
\end{equation}
The Lorenz curve maximum is the $\lfloor n/2 \rfloor$-th elbow and its coordinates are calculated by collecting all the positive Wigner components,
\begin{align}
	x_{\lfloor n/2 \rfloor} &= \frac{1}{2}\left(1 + \left(\frac{7}{9}\right)^n\right), \\
	L_{\lfloor n/2 \rfloor} &= \frac{1}{2}\left (1 + \left(\frac{15 - 8\epsilon}{9}\right)^n \right).
\end{align}
%\sum_{j: even}^n a^j \binom{n}{j} = \frac{1}{2} [ (1+a)^n + (1-a)^n ]

@@

Let each qutrit have a Hamiltonian $H$ with energies $E_0, E_1, E_2$ and stabilizer eigenstates, and define $E_{\rm max} = \max\{E_0, E_1, E_2\}$ and $E_s$ as the eigenvalue of the Hamiltonian eigenstate whose Wigner distribution overlaps the negative component of $\rho_S(\epsilon)$ in the phase space. 
Let $T =(k\beta)^{-1}$ be any finite temperature and assume that in the thermodynamic limit ($n,m \gg 1$), the protocol applied to the equilibrium state $\tau^{\otimes n} = (e^{-\beta H}/\Z)^{\otimes n}$ maps $\tau^{\otimes n} \longrightarrow \tau^{\prime \otimes m}$, where we express state $\tau'$ as $\tau' = e^{-\beta H'}/\Z'$ for some Hermitian $H'$.

@@

The equilibrium state at inverse temperature $\beta$ on a single qutrit is given by $\tau = e^{-\beta H} / \Z$. Moreover we have that $\tau$ is a full-rank stabilizer state, where $\beta \geq 0$ and $H = E_0 \ketbra{\varphi_0} + E_1 \ketbra{\varphi_1} + E_2 \ketbra{\varphi_2}$ is an eigendecomposition of $H$.
The state $\tau$ can now be written as 
\begin{equation}
	\tau = \frac{e^{-\beta E_0}}{\Z} \ketbra{\varphi_0} + \frac{e^{-\beta E_1}}{\Z} \ketbra{\varphi_1} + \frac{e^{-\beta E_2}}{\Z} \ketbra{\varphi_2},
\end{equation}
where the eigenstates $\{\ket{\varphi_k}\<\varphi_k|\}$ are pure, orthonormal stabiliser states, which can be represented in terms of generalized Paulis. We are free to redefine the computational basis so it coincides with the basis of $H$. Abstractly, let $C$ be the unitary transforming each $|\varphi_k\>\<\varphi_k|$ to $|k\>\<k|$. Since the Clifford group is the normalizer of the Heisenberg-Weyl group, $C$ is a Clifford unitary. Therefore, $C$ maps $\tau$ to another stabilizer state that is diagonal in the computational basis, and we can assume without loss of generality that $\tau$ is diagonal in $\{|0\>,|1\>, |2\>\}$. However, this re-definition of the computational basis means that the location of the negative Wigner component $-v(\epsilon)$ of the Strange state will be permuted on the discrete phase space. We denote by $E_s$ the eigenvalue of $H$ where the associated eigenvector has Wigner distribution overlapping the negative component of the magic state $C\rho_S(\epsilon)C^\dagger$.  This is unique, since the eigenstates form an orthonormal basis. \nick{Clifford processing?}

>>>>

If we have a magic state $\rho$ that has negativity in its Wigner representation, then, as discussed, its Lorenz curve over-shoots $1$ and reaches a non-trivial maximum $L_\star$ that depends on the particular state. There is an exact relation between $L_\star$ and sum-negativity or mana, which is provided by the following result. 
\begin{theorem}\label{lem:lcmax}
	Given a magic state $\rho$, the maximum $L_\star$ of its Lorenz curve $\lc{\rho}{\sigma}(x)$ is independent of the $\sigma$--fragment and equal to $1+\sn{\rho}$. Moreover, the majorization constraint is stronger than mana in every fragment.
\end{theorem}
The proof of this is given in~\cref{app:major}. Using the Lorenz curve perspective, it is simple to construct other monotones. For example, within a given fragment, the area above the horizontal line $y = 1$ is another magic monotone.
\begin{lemma}
Given a magic state $\rho$ and a free state $\sigma$, let $\A_\sigma(\rho)$ be the area of the region $\{(x, y): 1 \leq y \leq L_{\rho | \sigma}(x)\}$. Then $\A_\sigma(\rho)$ is a magic monotone for $\R_\sigma$.
\end{lemma}
\begin{proof}
Consider the transformation $\rho_1 \rightarrow \rho_2$ within $\R_\sigma$. We have that $L_{\rho_1|\sigma}(x)$ is never below the curve $L_{\rho_2|\sigma}(x)$, and therefore the region above $L=1$ for $\rho_2$ is a subset of the corresponding region for $\rho_1$. Thus, $\A_\sigma(\rho_1) \geq \A_\sigma(\rho_2)$, so $\A_\sigma$ is a magic monotone in $\R_\sigma$.
\end{proof}
In contrast to mana, the area monotone is \emph{specific to the fragment}, and its value will vary as we change $\sigma$. Therefore its monotonicity is also including the physics of the fragment and provides a means to analyse magic distillation only under free operations that leave $\sigma$ invariant.

@@

We now show that the Lorenz curve is concave in the context of quasi-distributions.
\begin{lemma}\label{L-concave} Let $\p$ be a quasi-probability distribution and let $\r$ be a probability distribution with strictly non-zero components. Then $L_{\p|\r}(x)$ is a concave function on $[0,1]$.
\nick{
\begin{proof}
	The Lorenz curve of $\p$ relative to $\r$ passes through the origin $(r_{\pi(0)}, p_{\pi(0)}) := (0,0)$ and the points $(\sum_{i=1}^k r_{\pi(i)}, \sum_{i=1}^k p_{\pi(i)})$ where $\pi$ is the permutation that puts $p_i/r_i$ in non-increasing order.
	By construction, the slope of the curve attains the values 
\begin{equation}
	\frac{\sum_{i=0}^k p_{\pi(i)} - \sum_{i=0}^{k-1} p_{\pi(i)}}{\sum_{i=0}^k r_{\pi(i)} - \sum_{i=0}^{k-1} r_{\pi(i)}} = \frac{p_{\pi(k)}}{r_{\pi(k)}},\ \text{with } k \geq 1,
\end{equation}
throughout the interval $[0,1]$, which is a non-increasing sequence due to the definition of $\pi$.
	Therefore, the Lorenz curve is concave.
\end{proof}
}
\begin{proof}Let $L_{\p|\r}(x)$ attain its maximum at $x=x_\star$. Then from the definition, the function is continuous, monotone increasing on $[0,x_\star]$, and monotone decreasing on $[x_\star, 1]$. Consider $a,b \in [0,1]$ with $a <b$ and some probability $q \in [0,1]$. If $a$ and $b$ both lie in either the interval $[0,x_\star]$ or $[x_\star, 1]$ then concavity follows immediately. Otherwise, $a$ is in the first interval and $b$ in the second interval. If $w=qa + (1-q)b$ lies in the first interval then
\begin{align}
qL_{\p|\r}(a) + (1-q)L_{\p|\r}(b) &\le qL_{\p|\r}(a) + (1-q)L_{\p|\r}(x_\star)\nonumber\\
& \le L_{\p|\r}(w)
\end{align}
\nick{second inequality isn't clear to me}
from monotonicity on the first interval. Otherwise, $w$ lies in the second interval and
\begin{align}
qL_{\p|\r}(a) + (1-q)L_{\p|\r}(b) &\le qL_{\p|\r}(x_\star) + (1-q)L_{\p|\r}(b)\nonumber\\
& \le L_{\p|\r}(w)
\end{align}
\nick{again second inequality isn't clear to me}
which follows from monotonicity on the second interval. Therefore the Lorenz curve is concave on $[0,1]$.
\end{proof}
\end{lemma}

@@

In the unital fragment, the rescaled distribution is simply proportional to the Strange state distribution,
\begin{equation}
	W_{\rho_{\rm{S}}(\epsilon)^{\otimes n}|\left(\id/3\right)^{\otimes n}} = 9^n W_{\rho_{\rm{S}}(\epsilon)^{\otimes n}},
\end{equation}
therefore expressions for the Lorenz curve coordinates follow from summing up the Wigner components in decreasing order.
\nick{As a result, every Strange state Lorenz curve in the unital fragment contains $n$ elbows. In the case of accidental degeneracies among Wigner components, some of these elbows may be straightened out.}
Including the boundary points $(x_{-1}, L_{-1}) := (0,0)$ and $(x_{n}, L_{n}) := (1,1)$, we label these elbows as 
\begin{equation*}
\{(x_{i}, L_{i})\}_{i=-1,0,\dots,n},
\end{equation*}
where the coordinates are given by
\begin{equation}
	(x_{i}, L_{i}) = \left( \frac{1}{9^n}\sum_{\ell=0}^i m_{\ell}, \sum_{\ell=0}^i m_{\ell} w_{\ell} \right).
\end{equation}
The Lorenz curve maximum is the $\lfloor n/2 \rfloor$-th elbow and its coordinates are calculated by collecting all the positive Wigner components,
\begin{align}
	x_{\lfloor n/2 \rfloor} &= \frac{1}{2}\left(1 + \left(\frac{7}{9}\right)^n\right), \\
	L_{\lfloor n/2 \rfloor} &= \frac{1}{2}\left (1 + \left(\frac{15 - 8\epsilon}{9}\right)^n \right).
\end{align}
%\sum_{j: even}^n a^j \binom{n}{j} = \frac{1}{2} [ (1+a)^n + (1-a)^n ]

@@

Let each qutrit have a Hamiltonian $H$ with energies $E_0, E_1, E_2$ and stabilizer eigenstates, and define $E_{\rm max} = \max\{E_0, E_1, E_2\}$ and \nick{$E_s$ as the eigenvalue of the Hamiltonian eigenstate whose Wigner distribution overlaps the negative component of $\rho_S(\epsilon)$ in the phase space.}
Let $T =(k\beta)^{-1}$ be any finite temperature and assume that in the thermodynamic limit ($n,m \gg 1$), the protocol applied to the equilibrium state $\tau^{\otimes n} = (e^{-\beta H}/\Z)^{\otimes n}$ maps $\tau^{\otimes n} \longrightarrow \tau^{\prime \otimes m}$, where we express state $\tau'$ as $\tau' = e^{-\beta H'}/\Z'$ for some Hermitian $H'$.

@@

\nick{Explain why this is not Clifford processing, ensure overlap with negativity is the one suggested}
The equilibrium state at inverse temperature $\beta$ on a single qutrit is given by $\tau = e^{-\beta H} / \Z$. Moreover we have that $\tau$ is a full-rank stabilizer state, where $\beta \geq 0$ and $H = E_0 \ketbra{\varphi_0} + E_1 \ketbra{\varphi_1} + E_2 \ketbra{\varphi_2}$ is an eigendecomposition of $H$.
The state $\tau$ can now be written as 
\begin{equation}
	\tau = \frac{e^{-\beta E_0}}{\Z} \ketbra{\varphi_0} + \frac{e^{-\beta E_1}}{\Z} \ketbra{\varphi_1} + \frac{e^{-\beta E_2}}{\Z} \ketbra{\varphi_2},
\end{equation}
where the eigenstates $\{\ket{\varphi_k}\<\varphi_k|\}$ are pure, orthonormal stabiliser states, which can be represented in terms of generalized Paulis. We are free to redefine the computational basis so it coincides with the basis of $H$. Abstractly, let $C$ be the unitary transforming each $|\varphi_k\>\<\varphi_k|$ to $|k\>\<k|$. Since the Clifford group is the normalizer of the Heisenberg-Weyl group, $C$ is a Clifford unitary. Therefore, $C$ maps $\tau$ to another stabilizer state that is diagonal in the computational basis, and we can assume without loss of generality that $\tau$ is diagonal in $\{|0\>,|1\>, |2\>\}$. However, this re-definition of the computational basis means that the location of the negative Wigner component $-v(\epsilon)$ of the Strange state will be permuted on the discrete phase space. We denote by $E_s$ the eigenvalue of $H$ where the associated eigenvector has Wigner distribution overlapping the negative component of the magic state $C\rho_S(\epsilon)C^\dagger$.  This is unique, since the eigenstates form an orthonormal basis.
